\newpage

\begin{center}
    {\large \bf Acknowledgements }
\end{center}

\vspace{-0.2cm}

First and foremost, I am privileged to have had Kathryn Johnston as my advisor
and mentor for the last four years. I am so grateful to have had an advisor who
supports me in my work and in life. Her intuition and ease at coming up with
creative new ideas is matched by her supportive, patient, and light-hearted
nature. Kathryn constantly (and accidentally) reminds me why I love doing
science and has been a huge influence on my [...]. I can't imagine a better
mentor---thanks for putting up with me!

Thanks to David Hogg for geting me interested in astronomy. I've always loved
science and have been fascinated by astronomy, but I didn't think I wanted to be
or could be a scientist. I started as an undergraduate in technical theater at
the Tisch School for the Arts at NYU. After one semester I knew that I wanted to
switch majors, but it wasn't until I took Physics 1 (taught by Hogg) that I knew
I wanted to do \emph{whatever that guy did}. Over the last ten years, he has
been an invaluable mentor, good influence, bad influence, and friend.

New York can be caustic, but I am lucky to have had friends, advocates, and
peers who have helped me tremendously throughout my life in the city. From the
scientific world, thank you Kelle Cruz, Dan D'Orazio, and Josh Peek for being
there. In the other world, huge thanks to Jared Hiller for putting up with me in
our (but really his) musical ventures. Thanks to my band, \emph{Operator Music
Band}, for getting me out of the office and letting me play music (Jared Hiller,
Dara Hirsch, Jon Young, and Landen Griffith). Thanks Adam, Peter, and Sean for
being there no matter what. Thanks also to the Ding Dong Lounge for the late
nights and grime.

I've received advice and guidance from so many people throughout graduate
school. Thanks espcially to (then) graduate students Jeff Andrews, Ana Bonaca,
Steph Douglas, Dan Foreman-Mackey, David Hendel, Cameron Hummels, Duane Lee,
Munier Salem, Jonathan Sick, but also to all of the Columbia graduate students
that I've been priveleged to know, chat with, play pool with, and learn from. To
officemates Lauren Corlies and Sarah Pearson. To postdocs Andy Casey, Andreas
K\"upper, Chervin Laporte, Dustin Lang, Melissa Ness, Robyn Sanderson, Branimir
Sesar, Allyson Sheffield, Nick Stone, Erik Tollerud, Matt Turk, Lucianne
Walkowicz. Thanks to faculty Marcel Ag\"ueros, Arlin Crotts, Julianne Dalcanton,
Marla Geha, Jacqueline van Gorkom, Zoltan Haiman, Jules Halpen, Jerry Ostriker,
and David Schiminovich for ideas, discussion, debates, and support, some of
which happen primarily through \project{Twitter}.

Special thanks to Hans-Walter Rix for supporting me in Germany each summer and
for welcoming me to his house and garden.

Huge thanks to Summer Ash for her friendship and for organizing outreach events
at Columbia. Herding cats is \emph{hard}!

Thanks to everyone in the Astropy community for letting me share in something
that has made me a better programmer and person. Especially Tom Robitaille,
Perry Greenfield, and Erik Tollerud (who gets a second thank you!). Thanks also
to the .astronomy community for the enthusiasm, crazy ideas, and fantastic
conferences. Especially Alyssa Goodman, Chris Lintott, and Rob Simpson.

I'm quite certain I would not have made it this far without the support and
assistance of Millie and Ayoune, the astronomy department staff. Thanks so much
for all of the help! Thanks also to the staff at MDM observatory, especially
Eric Galayda.

As an undergraduate, I had the pleasure of working with and learning from
Michael Blanton, Demitri Muna, and Ben Weaver.

I am fortunate to have lived so close to my family who have been there to [...]. Thanks to my parents, Audrey and Michael, who taught me to wonder. Thanks to my grandmother Pat and my sister Alexa and her amazing family, Lars and Seren, for the laughs, the [..], and the [...]. Also for letting me use her car!

Finally and most [...], thanks to Lauren for the love, adventures, patience, and for keeping me balanced over the last two and a half years. I can't imagine trying to do this without the love we have for each other.

\vspace{1.8cm}
2016, New York, NY

%\setlength{\baselineskip}{1.11111 \baselineskip}
